Algebraic File Synchronization: Adequacy and Completeness

### 2016-10-08 p2/works-inverse.tex
New notation for commands

\begin{myax}\label{combine_independent_commands}
The combination of independent commands works wherever the original commands work:
\[ \cxynv\indep \czwmv \Rightarrow \worksc{\cxynv\cc \czwmv}{\cxynv, \czwmv}. \]
% \begin{align*}
% \forall\FS:{}&\cxynv\indep \czwmv \\
% &\quad\wedge \cxynv\aFS\neq\fsbroken \\
% &\quad\wedge \czwmv\aFS\neq\fsbroken \\
% &\Rightarrow \cxynv\cc \czwmv\aFS\neq\fsbroken.
% \end{align*}
\end{myax}
\begin{proof}
We name this proposition a \namecref{combine_independent_commands} 
because to prove it, we must reach back to our filesystem model.
We proceed in an indirect way and
assume that there is a filesystem $\FS$ for which
$\cxynv\aFS\neq\fsbroken$ and $\czwmv\aFS\neq\fsbroken$, but
$\cxynv\cc \czwmv\aFS=\fsbroken$.
We know $\cxynv\aFS\neq\fsbroken$ so it must be applying 
$\czwmv$ that breaks it.
Applying a command can only result in a broken filesystem in three cases.
First, if the filesystem was already broken, which cannot be the case here.
Second, if the input type does not match the filesystem,
but we know $\FS(m)\in\setvx{W}$ and so
$(\cxynv\aFS)(m)\in\setvx{W}$ as based on \cref{incomparable_is_independent}, $n\neq m$.
Third, if the new filesystem violates the tree property.
This again cannot be the case because we also know that $n\unrel m$
and the tree property only depends on the types of the parent and children of $m$,
which therefore cannot be changed by $\cxynv$.
\end{proof}

@@

\begin{mylem}\label{worksinputmatch}
If $A$ and $B$ are minimal sequences, $\works{A,B}$,
and there are commands on node $n$ in both $A$ ($\cxynv\in A$) and $B$ ($\czwnv\in B$),
then the input types of these commands must match ($X=Z$).
\end{mylem}
\begin{proof}
This result is similar to \cref{equiv_simple_same_commands} and
is easily shown using an indirect proof: if $X\neq Z$, then there is no filesystem that
either $\cxynv$ or $\czwnv$ would not break, 
and consequently $A$ and $B$ cannot work on the same filesystem.
\end{proof}

@@

\noindent
We know that whether commands break a filesystem depends solely on their input and output
types and the node, but not their actual output value.
To represent this, we introduce an equivalence relation between filesystems, commands
and sequences of commands.
\begin{mydef}[Equivalence by type: $\typeeq$]
$ $ % otherwise first list element starts on same line
\begin{itemize}
\item For two filesystems we write $\FS\typeeq\GS$ iff they are both broken, or they share the same nodes,
and at all nodes the types of their values are the same.
\item For two commands we write $\alpha\typeeq\beta$ iff their input and output types,
and the node they operate on are the same.
For example, $\cbfaa{n}{f_1}\typeeq\cbfaa{n}{f_2}$.
\item For two sequences of commands we write $A\typeeq B$ iff they contain the same number of commands,
and pairwise the commands $\alpha\in A$ and $\beta\in B$ at the same indices are equivalent by type ($\alpha\typeeq\beta$).
\end{itemize}
\end{mydef}

In other words, two type-equivalent objects can only differ in the value they contain,
but as $|\setb|=|\setd|=1$, this is only possible at values of the type $\ccharf$.
We select an arbitrary file value $\fil\in\setf$, and choose a representative
of each $\typeeq$-equivalence class where all (if any) values of the type $\ccharf$ is $\fil$.
For example, in the case of commands, the representative of $[\cbfaa{n}{f}]_\typeeq$ is $\cbfaa{n}{\fil}$.

\begin{mydef}[$\T$]
We use $\T$ and write $x\T$ to denote the canonical mapping of 
filesystems, commands and sequences to the
representatives of their $\typeeq$-equivalence class.
$\T$ is also extended to sets of sequences,
where $\seqset{A}\T = \{S\T | S\in\seqset{A}\}$.
\end{mydef}

For example, $\cbfaa{n}{f}\T = \cbfaa{n}{\fil}$,
and for the sequence $S=\alpha\cc\beta$, $S\T=\alpha\T\cc\beta\T$.
In other words, $\T$ maps our system of filesystems and commands
into a system where $|\setf|$ is also $1$.

@@

\begin{mylem}
Representative functions $\alpha\T$ are injections over representative filesystems ($\FS\T$)
except for the cases where they break a filesystem, that is,
\[ \forall\FS,\GS,\alpha: \FS\T\neq\GS\T \wedge \alpha\T(\FS\T)\neq\fsbroken \Longrightarrow \alpha\T(\FS\T) \neq \alpha\T(\GS\T). \]
\end{mylem}
\begin{proof}
This is trivial if $\GS\T=\fsbroken$, so we assume it is not so.
If $\FS\T$ and $\GS\T$ have different values at any node apart from $n$, then this difference remains
even after applying the command to $\FS\T$. The only remaining case therefore is where they have the same values
at all nodes except $n$.
We know that a representative filesystem can only have values from the set $\rvalueset$.
As $\alpha\T(\FS\T)\neq\fsbroken$, $\FS\T(n)$ must be the value that matches the input type of $\alpha\T$.
We know that $\GS\T(n)\neq\FS\T(n)$, and therefore $\GS\T(n)$ must have a different type,
which implies $\alpha\T(\GS\T)=\fsbroken$ as it cannot match the input type of $\alpha\T$,
and again we have $\alpha\T(\FS\T) \neq \alpha\T(\GS\T)$.
\end{proof}

@@

>>>>

\begin{myax}\label{combine_independent_commands}
The combination of independent commands works wherever the original commands work:
\[ \alpha\indep \beta \Rightarrow \worksc{\alpha\cc \beta}{\alpha, \beta}. \]
% \begin{align*}
% \forall\FS:{}&\cxynv\indep \czwmv \\
% &\quad\wedge \cxynv\aFS\neq\fsbroken \\
% &\quad\wedge \czwmv\aFS\neq\fsbroken \\
% &\Rightarrow \cxynv\cc \czwmv\aFS\neq\fsbroken.
% \end{align*}
\end{myax}
\begin{proof}
We name this proposition a \namecref{combine_independent_commands} 
because to prove it, we must reach back to our filesystem model.
Let $\alpha=\cxynv$ and $\beta=\czwmv$.
%% TODO by contradiction
We proceed in an indirect way and
assume that there is a filesystem $\FS$ for which
$\cxynv\aFS\neq\fsbroken$ and $\czwmv\aFS\neq\fsbroken$, but
$\cxynv\cc \czwmv\aFS=\fsbroken$.
We know $\cxynv\aFS\neq\fsbroken$ so it must be applying 
$\czwmv$ that breaks it.
Applying a command can only result in a broken filesystem in three cases.
First, if the filesystem was already broken, which cannot be the case here.
Second, if the input type does not match the filesystem,
but we know $[\FS(m)]=[w]$ and so
$[(\cxynv\aFS)(m)]=[w]$ as based on \cref{incomparable_is_independent}, $n\neq m$.
Third, if the new filesystem violates the tree property.
This again cannot be the case because we also know that $n\unrel m$
and the tree property only depends on the types of the parent and children of $m$,
which therefore cannot be changed by $\cxynv$.
\end{proof}

@@

\begin{mylem}\label{worksinputmatch}
If $A$ and $B$ are minimal sequences, $\works{A,B}$,
and there are commands on node $n$ in both $A$ ($\cxynv\in A$) and $B$ ($\czwnv\in B$),
then the input types of these commands must match ($[x]=[z]$).
\end{mylem}
\begin{proof}
This result is similar to \cref{equiv_simple_same_commands} and
%% TODO by contradiction
is easily shown using an indirect proof: if $[x]\neq [z]$, then there is no filesystem that
either $\cxynv$ or $\czwnv$ would not break, 
and consequently $A$ and $B$ cannot work on the same filesystem.
\end{proof}

@@

\noindent
We know that whether commands break a filesystem depends solely on their input and output
types and the node, but not their actual output value.
To represent this, we introduce an equivalence relation between filesystems, commands
and sequences of commands.
%% TODO extending it!
\begin{mydef}[Equivalence by type: $\typeeq$]
$ $ % otherwise first list element starts on same line
\begin{itemize}
\item For two filesystems we write $\FS\typeeq\GS$ iff they are both broken, or they share the same nodes $\setn$,
and for all $n\in\setn$, $[\FS(n)]=[\GS(n)]$.
\item For two commands we write $\cxynv\typeeq\czwnv$ iff 
$[x]=[z]$ and $[y]=[w]$ and $n=m$.
\item For two sequences of commands we write $A\typeeq B$ iff they contain the same number of commands,
and pairwise the commands $\alpha\in A$ and $\beta\in B$ at the same indices are equivalent by type ($\alpha\typeeq\beta$).
\end{itemize}
\end{mydef}

In other words, two type-equivalent objects can only differ in the value they contain,
but as $|\setb|=|\setd|=1$, this is only possible at values of the type $[\valf]$.
We select an arbitrary file value $\fil\in[\valf]$, and choose a representative
of each $\typeeq$-equivalence class where all (if any) values of the type $[\valf]$ is $\fil$.
For example, in the case of commands, the representative of $[\cbfaa{n}]_\typeeq$ is $\caaa{\empt}{\fil}{n}$.

\begin{mydef}[$\T$]
We use $\T$ and write $X\T$ to denote the canonical mapping of 
filesystems, commands and sequences to the
representatives of their $\typeeq$-equivalence class.
$\T$ is also extended to sets of sequences,
where $\seqset{A}\T = \{S\T | S\in\seqset{A}\}$.
\end{mydef}

For example, $\cbfaa{n}\T = \caaa{\empt}{\fil}{n}$,
and for the sequence $S=\alpha\cc\beta$, $S\T=\alpha\T\cc\beta\T$.
In other words, $\T$ maps our system of filesystems and commands
into a system where $|\setf|$ is also $1$.

% TODO <[x],y,n> T = <[x],[y],n>
% one-to-one corresp btw values and types

@@

\begin{mylem}
Representative commands $\alpha\T$ are injections over representative filesystems ($\FS\T$)
except for the cases where they break a filesystem, that is,
\[ \forall\FS,\GS,\alpha: \FS\T\neq\GS\T \wedge \alpha\T(\FS\T)\neq\fsbroken \Longrightarrow \alpha\T(\FS\T) \neq \alpha\T(\GS\T). \]
\end{mylem}
\begin{proof}
This is trivial if $\GS\T=\fsbroken$, so we assume it is not so.
If $\FS\T$ and $\GS\T$ have different values at any node apart from $n$, then this difference remains
even after applying the command to $\FS\T$. The only remaining case therefore is where they have the same values
at all nodes except $n$.
We know that a representative filesystem can only have values from the set $\rvalueset$.
As $\alpha\T(\FS\T)\neq\fsbroken$, $\FS\T(n)$ must be the value that matches the input type of $\alpha\T$.
We know that $\GS\T(n)\neq\FS\T(n)$, and therefore $\GS\T(n)$ must have a different type,
which implies $\alpha\T(\GS\T)=\fsbroken$ as it cannot match the input type of $\alpha\T$,
and again we have $\alpha\T(\FS\T) \neq \alpha\T(\GS\T)$.
\end{proof}

@@

%% TODO
% Do we need <[x],[y],n>-1 = <[y],[x],n> ?
% Do we need to prove if a filesystem is not in the image, then a-1 will break it?
